### 2019-02-18 README.Rmd
initial commit to euctrl-pru repo

---
output: github_document
---

This is the repo for the Reproducible ANS Performance paper.

You find here all the instructions to
fully reproduce the paper: the raw and pre-processed data, the code for the analysis and the
text of the paper. The latter two come in the form of an
[RMarkdown](https://rmarkdown.rstudio.com/) file `paper.Rmd`
that embeds prose and code chunks in the form of a 
[literate programming](https://en.wikipedia.org/wiki/Literate_programming) document.

Visit the [companion web site](https://euctrl-pru.github.io/reproducible-ans-performance-paper/)
for further details.