### 2019-02-18 analysis/data.Rmd

---
title: "Data"
output:
  workflowr::wflow_html:
    toc: false
---

Here we describe the
* raw data, i.e. the Reference Trajectory data set
* pre-processed data, i.e. the extracted arrivals and the transition points.