### 2019-02-18 analysis/index.Rmd

---
title: "Home"
output:
  workflowr::wflow_html:
    toc: false
---

This webpage describes the GitHub repository hosting the text, code and data for
the paper on Reproducible ANS Performance.

The aim of this repository is to make it possible to fully reproduce our analysis
and results.