### 2019-02-18 biographies.tex

\setlength{\parskip}{6pt}

\vspace*{1\baselineskip}
\hypertarget{biographies}{%
\section*{Authors' Biographies}\label{biographies}}
\addcontentsline{toc}{section}{Authors' Biographies}

\vspace*{-3\baselineskip}

\begin{IEEEbiographynophoto}{\normalsize Enrico Spinielli}\normalsize
is the Data Manager of the Performance Review Unit at EUROCONTROL, Brussels.
He holds a Master Degree in Electronic Engineering from the University of Pisa, Italy.
His intellectual interests include data analytics, data visualization and
map projections.
\end{IEEEbiographynophoto}

\vspace*{-2.5\baselineskip}

\begin{IEEEbiographynophoto}{\normalsize Rainer Koelle}\normalsize
heads the Safety and Quality of Service section
of the Performance Review Unit at EUROCONTROL, Brussels.
He holds a PhD from the University of Lancaster, 2013, in Communication Systems.
His research interests include data analytics for operational performance analysis.
\end{IEEEbiographynophoto}

\vfill