### 2019-02-18 exploration/check-coverage.Rmd

---
title: "Checking Completeness"
author: "RQ & ES"
date: "4 February 2019"
output: html_document
---

```{r, message=FALSE}
library(readr)
library(dplyr)
library(lubridate)
library(purrr)
library(ggplot2)
```

## Overview

This RMarkdown document is a quick hack to check the coverage of the ref trajectory data vs APDF.

## Read in APDF Data

For this paper we prepared a data set containing the arrivals at London Heathrow and Dublin (and Zurich - tbd)

```{r}
egll2017 <- read_csv("../data/egll_arrs_fact_2017.csv")
eidw2017 <- read_csv("../data/eidw_arrs_fact_2017.csv")
```

```{r}
start_date = lubridate::dmy_hms("01-08-2017 00:00:00", tz = "UTC")
end_date   = lubridate::dmy_hms("08-08-2017 23:59:59", tz = "UTC")

trim_time_horizon <- function(arrs, .start_date = start_date, .end_date = end_date){
  arrs <- arrs %>% mutate(ARR_TIME = ifelse(is.na(AIBT), STA, AIBT)) %>%
    filter(ARR_TIME >= start_date, ARR_TIME <= end_date) %>%
    select(-ARR_TIME)
  return(arrs)
}
```

## Read In Trajectory Flight Data

Trajectory derived flight data is stored in daily samples.

```{r}
read_flight_trj <- function(apt, .path = "../data/"){
  data_path <- .path
  data_files<- dir(data_path, pattern = paste(apt, "_flt_rt", sep = ""))

  flts <- data_files %>%
    map(~ read_csv(file.path(data_path, .),
      col_types = cols(
         period_start = col_datetime(format = "")
        ,period_finish = col_datetime(format = "")
        ))
      ) %>%
    bind_rows()
  return(flts)
}
```

```{r, warning=FALSE}
egll2017_flt_trj <- read_flight_trj("egll")
```

## Plot Coverage

```{r}
extract_arrs_fm_trj <- function(ds) {
  flt_trj <- ds %>% 
    select(REG = aircraft_reg, TYP = aircraft_type, ADEP = adep, ADES = ades
           , period_start, period_finish) %>%
    mutate(ARR_TIME = if_else(!is.na(period_finish), period_finish, period_start)
          # ,ARR_TIME = as.POSIXct(ARR_TIME, origin = "1970-01-01")
           ) %>%
    select(-contains("period"))
}
```

```{r, warning=FALSE}
egll_arrs_trj <- read_flight_trj("egll") %>% 
  extract_arrs_fm_trj() %>%
  mutate(DOF = lubridate::date(ARR_TIME), SOURCE = "TRJ") %>%
  filter(ARR_TIME <= end_date)

egll_arrs_apdf <- egll2017 %>% trim_time_horizon() %>% 
  mutate(ARR_TIME = if_else(!is.na(AIBT), AIBT, STA) 
         , DOF = lubridate::date(ARR_TIME),SOURCE = "APDF")%>% 
  select(REG, ADEP, ADES, ARR_TIME, DOF, SOURCE)
```

```{r}
tmp <- bind_rows( egll_arrs_trj, egll_arrs_apdf)

ggplot(tmp, mapping = aes(x = factor(DOF), group = SOURCE, fill = SOURCE)) + geom_bar(position = "dodge")
```

```{r}
egll_arrs_trj %>% group_by(REG, DOF) %>% summarise(N = n()) %>% arrange(desc(N))
```

```{r}
tmp <- bind_rows( egll_arrs_trj %>% na.omit, egll_arrs_apdf)

ggplot(tmp, mapping = aes(x = factor(DOF), group = SOURCE, fill = SOURCE)) + geom_bar(position = "dodge")
```

# Dublin EIDW

```{r, warning=FALSE}
eidw_arrs_trj <- read_flight_trj("eidw") %>% 
  extract_arrs_fm_trj() %>%
  mutate(DOF = lubridate::date(ARR_TIME), SOURCE = "TRJ") %>%
  filter(ARR_TIME <= end_date)

eidw_arrs_apdf <-  read_csv("../data/eidw_arrs_fact_2017.csv") %>% trim_time_horizon() %>% 
  mutate(ARR_TIME = if_else(!is.na(AIBT), AIBT, STA) 
         , DOF = lubridate::date(ARR_TIME),SOURCE = "APDF")%>% 
  select(REG, ADEP, ADES, ARR_TIME, DOF, SOURCE)
```

```{r}
tmp <- bind_rows( eidw_arrs_trj %>% na.omit, eidw_arrs_apdf)

ggplot(tmp, mapping = aes(x = factor(DOF), group = SOURCE, fill = SOURCE)) + 
  geom_bar(position = "dodge")
```

## Export

The aforementioned data processing is now wrapped into utility functions and ported to asma.R.
```{r}
rm(list = ls())  # reset to empty working environment
source("../R/asma.R")
```

```{r, warning=FALSE}
# load data sets
egll_arrs_apdf <- read_arrs_apdf("egll", .path = "../data/")
egll_arrs_trj  <- read_flts_trj( "egll", .path = "../data/")

# prepare coverage data set
egll_cov_apdf <- egll_arrs_apdf %>% prep_cov_apdf()
egll_cov_trj  <- egll_arrs_trj %>% prep_cov_trj()
```

```{r}
tmp <- bind_rows(egll_cov_apdf, egll_cov_trj) %>%
         trim_time_horizon %>% na.omit

ggplot(data = tmp
       ,mapping = aes(x = factor(DOF), fill = SOURCE)
       ) +
  geom_bar(position = "dodge") + 
  theme_bw()
```
```{r}
cov_tbl <- tmp %>% group_by(SOURCE, DOF) %>%
  summarise(N_ARR = n()) %>%
  tidyr::spread(SOURCE, N_ARR) %>%
  mutate(COV = (RTRJ - APDF)/APDF)
```
```{r}
knitr::kable(cov_tbl)
```