### 2019-02-18 exploration/prepare_ASMA_rwy.Rmd

---
title: "Untitled"
author: "RQ"
date: "11 February 2019"
output: html_document
---

Quick hack to augment study with a heuristic approach to identifying landing runway.

```{r}
library(tidyverse)
source("../R/asma.R")
#source("../R/egll-data.R")  # throws: Error in library(trrrj) : there is no package called ‘trrrj’
```
```{r}
# paste and copy EGLL data
egll_apt <- tibble::tribble(
     ~latitude,  ~longitude, ~elevation,  ~icao,          ~id, ~type,        ~name,
  "N51°28.65'", "W0°27.68'",       83.0, "EGLL",       "EGLL", "ARP",   "EGLL ARP",
  "N51°28.65'", "W0°29.10'",         79, "EGLL", "RWY09L/27R", "RWY", "Runway 09L",
  "N51°28.66'", "W0°26.00'",         78, "EGLL", "RWY09L/27R", "RWY", "Runway 27R",
  "N51°27.89'", "W0°28.94'",         75, "EGLL", "RWY09R/27L", "RWY", "Runway 09R",
  "N51°27.90'", "W0°26.04'",         77, "EGLL", "RWY09R/27L", "RWY", "Runway 27L"
)
```

## damn ... in degree, minute decimal notation ==> do it by hand!

```{r, message=FALSE}
egll_pts <- read_pts_trj("egll", .path = "../data/") %>%
  filter(type %in% c("P40", "PLAND"))
```

```{r}
ggplot(data = egll_pts, mapping = aes(x = latitude, y = longitude, colour = type)) +
  geom_point()
```

```{r}
tmp <- egll_pts %>% filter(type == "PLAND", distance_arp <= 5) %>%
  mutate(RWY_BRG = cut(bearing, breaks = c(0,58, 180, 280, 360)))

# note: bearings are from point to ARP!
# 0,58 --> 09R
#(58,180] -->09L
#(180,280] --> 27R
#(280,360] --> 27L

egll_arp <- data.frame(longitude = -0.461389, latitude = 51.4775, type = "ARP", RWY_BRG = NA)

ggplot(data = tmp , mapping = aes(x = longitude, y = latitude, colour = RWY_BRG)) +
  geom_point() +
  geom_point(data = egll_arp, mapping = aes(size = 6))
```

What is wrong with the bearings for ASMA sectors?

```{r, message=FALSE}
egll_pts <- read_pts_trj("egll", .path = "../data/") %>%
  filter(type %in% c("P40", "PLAND"))
```

```{r}
my_ids <- c(egll_pts$flight_id[1], egll_pts$flight_id[17], egll_pts$flight_id[1723])

ggplot(data = egll_pts %>% filter(flight_id %in% my_ids)
       ,mapping = aes(x = longitude, y = latitude, colour = type)) +
  geom_point() + 
  geom_text(aes(label=bearing),hjust=0, vjust=0) +
  geom_text(aes(label=ifelse(bearing < 180, bearing + 180, bearing - 180)),hjust = -0.1 , vjust = 0.9, colour = "blue")
  ## ARP
  geom_point(data = egll_arp)

# bearing TO STATION == QDM !!!
```

Apply heuristic to EIDW

```{r, message=FALSE}
eidw_pts <- read_pts_trj("eidw", .path = "../data/") %>%
  filter(type %in% c("P40", "PLAND"))
```

```{r}
ggplot(data = eidw_pts, mapping = aes(x = latitude, y = longitude, colour = type)) +
  geom_point()
```

```{r}
# EIDW ARP
arp <- data.frame(longitude = -6.27, latitude = 53.421389, type = "ARP", RWY_BRG = NA)

# note: bearings are from point to ARP!
# 0,160 --> 10
#(160,210] --> 16
#(210,260] --> 34
#(260,360] --> 28

tmp <- eidw_pts %>% filter(type == "PLAND", distance_arp <= 5) %>%
  mutate(RWY_BRG = cut(bearing, breaks = c(0,160,210,260, 360)))

ggplot(data = tmp , mapping = aes(x = longitude, y = latitude
                                  , colour = RWY_BRG
                                  )   ) +
  geom_point() +
  geom_point(data = arp, mapping = aes(size = 6), colour = "white")
```