### 2019-02-18 exploration/replicate-ASMA.Rmd

---
title: "Replicate ASMA"
author: "RQ"
date: "3 January 2019"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

## Introduction

This R Markdown document is an initial stab at replicating the ASMA algorithm on the basis of the data extracted for this study/paper (i.e. ATM Seminar 2019).
The paper follows the concept of "teaching the computer to do the work". This way a track record of the data manipulation is achieved. In particular the first chapter is based on establishing the analytical data for ASMA. For the paper a subset of arrival flights are extracted from the FACT data tables (i.e. July 2017 for EGLL, EIDW, and LSZH).
The following sections work with this analytic data and serves as a validation of the trajectory derived data.

## Airport Operator Data Flow & 40NM data - reference data samples

For the determination of the ASMA indicator, the airport operator data flow (providing relevant flight-by-flight movement data) is augmented with a 40NM crossing 4-D position derived from the NM trajectory (i.e. profile 3).
The associated data is stored in the FACT table in PRISME.
The following extracts the relevant validation information from the FACT table.

With a view to establish a fully reproducible data sample, the respective FACT table data extraction is provided as "raw" data.

Note: The locally stored files have been retrieved in early 2018 as part of the international benchmarking. This is outside the scope of this paper.

```{r}
# utility functions to extract data for ASMA validation data set from FACT tables

# define required data items from FACT table
# Note: datetimes are not consistently stored in PRISME (some / or -). Thus, we 
# avoid throwing an error by defining time format with the any-separator, i.e. '%.'.
data_cols <- cols_only(   # begin pre-defined colums for study
       AP_C_FLTID = col_character()
      ,AP_C_REG   = col_character()
      ,AP_C_ARCTYP= col_character()
      ,AC_CLASS   = col_character()
      ,ADEP_ICAO  = col_character()
      ,ADES_ICAO  = col_character()
      ,SRC_PHASE  = col_character()
      ,SCHED_TIME_UTC = col_datetime(format = "%d%.%m%.%Y %H:%M:%S")
      ,BLOCK_TIME_UTC = col_datetime(format = "%d%.%m%.%Y %H:%M:%S")
      ,MVT_TIME_UTC  = col_datetime(format = "%d%.%m%.%Y %H:%M:%S")
      ,AP_C_STND  = col_character()
      ,AP_C_RWY   = col_character()
      ,C40_CROSS_TIME = col_datetime(format = "%d%.%m%.%Y %H:%M:%S")
      ,C40_CROSS_LAT  = col_double()
      ,C40_CROSS_LON  = col_double()
      ,C40_CROSS_FL  = col_integer()
      ,C100_CROSS_TIME = col_datetime(format = "%d%.%m%.%Y %H:%M:%S")
      ,C100_CROSS_LAT  = col_double()
      ,C100_CROSS_LON  = col_double()
      ,C100_CROSS_FL  = col_integer()
) # end - pre-defined columns

#' read locally stored FACT table
#'
#' FACT tables have been extracted from PRISME and stored locally in a
#' subfolder for all R projects, i.e. __DATA/. In this folder, there is
#' a subfolder per airport (e.g. EGLL/) with the associated FACT table
#' for the year.
#'
#' @param .apt 4-digit ICAO location indicator (chr string), e.g. "EGLL"
#' @param .yr  year (int), e.g. 2017
#' @param .path_to_fact (chr string) for folder & hierarchy
#'
#' @return filename (incldg. path)
#'
fact_file_name <- function(.apt, .yr, .path_to_fact){
  fn <- paste(.path_to_fact,"/", .apt,"/", .apt, "_", .yr, "_FACT.txt", sep = "")
  return(fn)
}

#' load extracted FACT table with predefined columns (data items)
#' 
#' Read in the specified set of data items from extracted FACT tables.
#'
#' @param .file_name (path and) filename of file to read
#' @param .data_cols cols_only() specification for readr::read_csv
#'
#' @return tibble of FACT table
#'
load_fact_data <- function(.file_name, .data_cols){
  ds <- readr::read_csv2(.file_name, col_types = .data_cols)
  return(ds)
}

#' load study data
#'
#' wrapper for defining file name and load of fact table
#'
load_data <- function(apt, yr, data_cols, path_to_fact = "../../__DATA/"){
  fn <- fact_file_name(.apt = apt, .yr = yr, .path_to_fact = path_to_fact)
  ds <- load_fact_data(.file_name = fn, .data_cols = data_cols)
  return(ds)
}

#' prepare loaded data for study
#'
#' The loaded data is trimmed for arrivals and the variables get
#' nicer names
trim_arrs <- function(ds){
  ds <- ds %>% dpyr::filter(SRC_PHASE == "ARR") %>%
    dplyr::select( CALLSIGN = AP_C_FLTID
           ,REG      = AP_C_REG
           ,ADEP     = ADEP_ICAO
           ,ADES     = ADES_ICAO
           ,ALDT     = MVT_TIME_UTC
           ,AIBT     = BLOCK_TIME_UTC
           ,STA      = SCHED_TIME_UTC
           ,TYP      = AP_C_ARCTYP
           ,WTC      = AC_CLASS
           ,C40_TIME = C40_CROSS_TIME
           ,C40_LAT  = C40_CROSS_LAT
           ,C40_LON  = C40_CROSS_LON
           ,C40_ALT  = C40_CROSS_FL
           )
  return(ds)
}

limit_time_horizon <- function(arrs, start = "01-07-2017", end= "31-08-2017"){
  start <- paste(start, "00:00:00", sep = "")
  start <- lubridate::dmy_hms(start, tz = "UTC")
  end   <- paste(end, "23:59:59", sep = "")
  end   <- lubridate::dmy_hms(end, tz = "UTC")

  arrs  <- arrs %>% 
    mutate(HORIZON = ifelse(is.na(AIBT), STA, AIBT)) %>%
    filter(HORIZON >= start, HORIZON <= end) %>%
    select(-HORIZON)

  return(arrs)
}

```

The following checks for the existance of the validation data sets.
If the files do not exist, the FACT table will be read and subsetted.
If the file exists, it will be loaded.

```{r, message=FALSE}
if(!file.exists("../data/egll_arrs_fact_2017.csv")){
  egll2017 <- load_data("EGLL", 2017, data_cols) %>% 
    trim_arrs() %>% 
    limit_time_horizon()
  write_csv(egll2017, "../data/egll_arrs_fact_2017.csv")
}else{
  egll2017 <- readr::read_csv("../data/egll_arrs_fact_2017.csv")
}

if(!file.exists("../data/eidw_arrs_fact_2017.csv")){
  eidw2017 <- load_data("EIDW", 2017, data_cols) %>% 
    trim_arrs() %>% 
    limit_time_horizon()
  write_csv(eidw2017, "../data/eidw_arrs_fact_2017.csv")
}else{
  eidw2017 <- readr::read_csv("../data/eidw_arrs_fact_2017.csv")
}

if(!file.exists("../data/lszh_arrs_fact_2017.csv")){
  lszh2017 <- load_data("LSZH", 2017, data_cols) %>% 
    trim_arrs() %>% 
    limit_time_horizon()
  write_csv(lszh2017, "../data/lszh_arrs_fact_2017.csv")
}else{
  lszh2017 <- readr::read_csv("../data/lszh_arrs_fact_2017.csv")
}
```

## Read In Trajectory Derived Data

Based on the trajectory derived data, we construct a "quick-and-dirty" ASMA data set.

```
To-do: identification of landing runway
--> we know 'landing end' of trajectory and should be able to determine distance and offset.
--> (alternatively, we can "steal" the runway from the FACT table)

To-do: enable read-out of xxxx-data.R without package
```

```{r, message=FALSE}
path_data <- "../data/" 
flt_fn    <- "egll_flt_rt_2017-08-01.csv"
pos_fn    <- "egll_pos_rt_40NM_2017-08-01.csv"
## need to load library trrrj
## source("../R/egll-data.R")
## EGLL ARP := "N51°28.65'", "W0°27.68'"
## converted with: https://www.directionsmag.com/site/latlong-converter/
egll_arp = data.frame(APT = "EGLL", PTS = "ARP", LAT = 51.4775, LON = 0.4613333)

flts <- readr::read_csv(paste(path_data, flt_fn, sep = ""))
poss <- readr::read_csv(paste(path_data, pos_fn, sep = ""))

flts <- left_join(flts, poss, by = "flight_id" ) %>% 
  select(flight_id, CALLSIGN = callsign, REG = aircraft_reg
         , TYP = aircraft_type, ADEP = adep, ADES = ades
         , TIME = timestamp, LAT = latitude, LON = longitude
         , DIST = distance  # is this the flown distance?
         , DIST_ARP = distance_arp 
         ) %>% 
  mutate(DIST_ARP_NM = DIST_ARP / 1852) %>% # convert to NM
  filter(DIST_ARP_NM >= 6)   # approx. short final = Gate ~ ALDT

asma_flts <- flts %>% group_by(flight_id, CALLSIGN) %>%
  summarise( C40_DIST = max(DIST_ARP_NM), C40_TIME = min(TIME)
            ,GATE_DIST= min(DIST_ARP_NM), GATE_TIME= max(TIME)
            ) %>% ungroup() %>%
  left_join(flts, by=c("flight_id", "CALLSIGN", "C40_TIME" = "TIME")
            ) %>% 
  mutate(BRG = geosphere::bearing(c(egll_arp$LON, egll_arp$LAT), cbind(LON, LAT))
         ,C40_BRG = (BRG + 360) %% 360 ) # modulo operation for 360-bearings

```

In order to determine the arrival sector for ASMA, we plot, the C40_BRG:

```{r}
ggplot(data = asma_flts, mapping = aes(x = C40_BRG)) + 
  geom_histogram(binwidth = 2)
```

Next: 
1. determine entry sectors
2. calculate travel time per entry sector and type
3. determine refernce time based on aforemntioned groups (still we need runway)