### 2019-02-18 paper.tex

%% template.tex
%% from
%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference,final,a4paper,]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%

%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

%% BEGIN MY ADDITIONS %%

\usepackage{longtable,booktabs}

\usepackage[unicode=true]{hyperref}

\hypersetup{
            pdftitle={Busting the Myth! - Reproducibility in ANS Performance},
            pdfauthor={Enrico Spinielli, Rainer Koelle},
            pdfkeywords={reproducibility, ANS performance, flight trajectory, open data, performance indicator},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{5}

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

% Pandoc header
%\def\tightlist{}

% \def\tightlist{%
%   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% \renewcommand*{\bibfont}{9pt}

% \usepackage[utf8x]{inputenc}
% \usepackage{mathptmx}

% \usepackage[T1]{fontenc}

% section title small caps issue
% Uncomment and knit once in order to force use of Times font
% as per IEEE Trans style. It can then stay commented out.
% \usepackage{times}

\usepackage{breakurl}
\usepackage{graphicx}

%% DO NOT use ULEM: it converts italics to underlined
% \usepackage{ulem} % ulem is needed to support strikethroughs (\sout) [with pandoc extension]

% \usepackage{biblatex}
\usepackage{doi}

\usepackage{float}

% as requested by conference Editor (plus hack after References section in Rmd)
\usepackage[skip=6pt,indent]{parskip}

% enable intentionally default disabled IEEE commands: needed for biographies
\IEEEoverridecommandlockouts

\def\IEEEkeywordsname{Keywords}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%% END MY ADDITIONS %%

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Busting the Myth! - Reproducibility in ANS Performance}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations

\author{

%% ---- classic IEEETrans wide authors' list ----------------
 % -- end affiliation.wide
%% ----------------------------------------------------------

%% ---- classic IEEETrans one column per institution --------
 %% -- beg if/affiliation.institution-columnar
\IEEEauthorblockN{
  %% -- beg for/affiliation.institution.author
Enrico Spinielli,
  %% -- beg for/affiliation.institution.author
Rainer Koelle %% -- end for/affiliation.institution.author
}
\IEEEauthorblockA{Performance Review Unit\\
EUROCONTROL\\
Brussels, Belgium
  %% -- beg for/affiliation.institution.author
\\enrico.spinielli@eurocontrol.int
  %% -- beg for/affiliation.institution.author
\\rainer.koelle@eurocontrol.int
 %% -- end for/affiliation.institution.author
}
 %% -- end for/affiliation.institution
 %% -- end if/affiliation.institution-columnar
%% ----------------------------------------------------------

%% ---- one column per author, classic/default IEEETrans ----
 %% -- end if/affiliation.institution-columnar
%% ----------------------------------------------------------

}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Air transportation is undergoing a fundamental transformation. On the
political, strategic, and operational level work is directed to meet the
future growth of air traffic. To meet this goal, operational concepts
and technological enablers are deployed with the promises of distinct
performance improvements. These improvements are typically widely
marketed, however, the underlying evidence is not made available. From
that respect Air Navigation System Performance suffers from the same
credibility crisis seen in other industries. This paper implements the
reproducibility paradigm by providing data, code, and results openly to
enable another researcher or analyst to reproduce - and potentially -
improve or build on our work. This work is part of a wider thread to
establish Open Data and Open Source based data analytical capabilities
for Air Navigation System Performance in Europe. The implementation of
the paradigm is demonstrated on the basis of the performance reference
trajectory by replicating a classical performance measure, and then
moving into investigating the application of trajectory based
information for enhancing the current state of performance monitoring
for the arrival phase. This paper applies the concept and approach as an
use case analysis for 2 European airports with different strategies for
sequencing arrival traffic (i.e.~stack holding and point merge). The
work reported demonstrates the feasibility of the approach and reflects
a stepping stone in the roll-out of the performance reference trajectory
within the European context. The work also builds the basis for novel
developments on the basis of the implemented reproducibility workflow as
this paper is accompanied by a webpage that supports the impartial
evaluation of the work by providing the data, code, and how this reports
had been produced.
\end{abstract}

% keywords
\begin{IEEEkeywords}
\textit{%
reproducibility; ANS performance; flight trajectory; open data; performance indicator}
\end{IEEEkeywords}

% use for special paper notices

% make the title area
\maketitle

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}\label{introduction}

Air transportation is undergoing a fundamental transformation. To meet
the future growth of air traffic, political decision makers, strategic
planers, practitioners and researchers aim to enhance the performance of
air navigation services (ANS). At the time being there is no consistent
and regionally harmonised deployment programme. Typically, the
deployment programmes are revolving around public-private partnerships
and regional funding programmes resulting in a wide mix of consortia and
stakeholder subsets. The actual deployment of novel concepts or
technological enablers is then driven by the programmes and to a lesser
extent by operational network wide priorities following a piece meal
approach.

Moreover, the success of these deployments is typically supported by
promising business cases and operational evaluations - often expressed
as operational performance improvements. While these improvements are
marketed, the underlying data and associated analyses in support of the
results are seldomly made available. Explanations range from data
sensitivity, commercial interests and property rights, through to
complexity of the data analyses or simulations, and the associated
volume of data.

Accordingly, ANS performance improvements are subject to a ``credibility
crisis'' similar to reported problems in other domains, e.g.~medicine
{[}\protect\hyperlink{ref-kolata_2011}{1}{]}, psychology
{[}\protect\hyperlink{ref-ritchie_2012}{2}{]}, ecology
{[}\protect\hyperlink{ref-fraser_etal_2018}{3}{]}. Fidler and Wilcox
(2018) provides a recent review of the reproducibility discussion
{[}\protect\hyperlink{ref-fidler_wilcox_2018}{4}{]}. Part of the crisis
sensation was driven by the diffusion of reported results as other
researchers/analysts were able to discover substantial flaws in the data
analyses. However, a crisis also offers room for self-correction.
Throughout the recent years, this lead to revisitng the concept of
reproducible research. This concept is based on the premises that an
independent researcher/analyst is enabled to reproduce the reported
results. Reproducibility covers therefore a spectrum and mix of method,
data, and analytical code/scripts per se openly available for scrutiny
by others.

Independent (operational) ANS performance builds on transparent and
robust performance monitoring and reporting. Within the European context
EUROCONTROL's Performance Review Unit (PRU) is committed to provide
impartial analyses of the performance of the European Air Navigation
System based on verified data. With a view to enhance transparency, PRU
is - inter alia - making performance-related data publicly available via
its Portal, \url{http://ansperformance.eu}, since early 2017 combined
with the methodological documentation for the calculation of the
performance indicators used under the EUROCONTROL Performance Revew
System and the European Union Performance Scheme. There has been
recognition of using common and openly available tools and data in ATM
research and operational validation. With the advent of crowd sourced
data and aviation databases there is an opportunity to advance the state
of the art by combining open data and government organisation collected
data to establish a curated platform supporting the quest for
reproducibility in ANS Performance. One of the associated building
blocks is the performance reference trajectory
{[}\protect\hyperlink{ref-spinielli_2018}{5}{]}. The performance
reference trajectory will be deployed as Open Data throughout 2019. The
trajectory project builds on crowd sourced surveillance data fused with
correlated position reports collected from European air navigation
service providers. Based on the Open Data model it will be possible to
augment the data processing (e.g.~data coverage, time synchronisation),
establishment of derived data bases (e,g fleet data base), associated
performance models (e.g.~aircraft performance model), but also the
continual further development and validation of operational performance
measures/indicators.

This paper builds on the previous work and implements the
reproducibility paradigm. As a use-case example, a fully reproducible
paper is devised assessing the performance in the arrival phase for 2
European airports. The underlying data, supporting data transformation
and analysis, and the paper ``source text'' are made openly available at
the paper companion web repository
\url{https://github.com/euctrl-pru/reproducible-ans-performance-paper}.
This enables the independent review and validation of the work reported.
Conceptually, an interested researcher/analyst can reproduce this paper
by using the same toolset or validate the results by using the source
data or interim data products if the complete reproduction of the study
would be too time/resource consuming.

The contributions of this paper are as follows:

\begin{itemize}
\tightlist
\item
  first fully reproducible ANS Performance related paper: The paper is
  written in RMarkdown allowing for the analysis of how text, code
  chunks, and visualisations are produced (including associated
  parameter settings) on the basis of the ``source text''.
\item
  transparent data analyses and result generation through supporting
  web-page: All data used in this paper, intermediate data artifacts,
  and results including the associated transformation are publicly
  available for inspection, validation, or reproduction.
\item
  use-case analysis of ANS Performance during the arrival phase for 2
  European airports: As an example the classical `additional ASMA time'
  performance indicator is implemented on the basis of the provided data
  with showcasing trajectory derived capabilities to further study the
  efficiency of the arrival phase (e.g.~impact of sequencing strategies
  {[}stack holdings, point merge{]}).
\end{itemize}

\section{Background -
Reproducibility}\label{background---reproducibility}

Conceptually, reproducibility is not a new topic. It is as old as the
scientific method and at the heart of the verification/validation of
results. The consistent and robust interpretation of results by
independent researchers/analysts forms the evidence and the inherent
body of knowledge in support of the applied methods or conclusions. This
principle is also relevant for the further development on the basis of
such results.

The terms replication and reproducibility are often used interchangeably
and there is a debate about the necessary and sufficient conditions for
these terms. Results are considered replicable, ``if there is sufficient
information available for independent researchers to make the same
{[}consistent{]} findings using the same procedures with new data''
({[}\protect\hyperlink{ref-gandrud_2015}{6}{]}, p.~3f). Replicability,
thus, does not aim to produce the ``same'' (aka identical) results. Here
the goal is to establish confirmation of the approach and findings by
deriving same conclusions.\footnote{For example, the same data
  collection procedures may not necessarily result in ``identical''
  results as the time horizon might be different, (new or different)
  data for another study case show different result values, but the
  conclusions are consistent with the replicated approach.} Accordingly,
results are reproducible, if an independent researcher can recreate the
findings based on the given information. In our domain (i.e.~operational
ANS Performance and associated data analysis) this will require the
access to data and derived summaries, how associated visualisations have
been generated, the accompanying explanatory text, and conclusions drawn
for the analysis.

Reproducibility has also to be seen in light of today's ``publish or
perish'' context. It appears that higher emphasis is given to the
frequency of publication than the validity of the produced results.
Within the medical world cases like the Duke ovarian cancer scandal
{[}\protect\hyperlink{ref-kolata_2011}{1}{]} are not single incidents,
but also demonstrate how errors during the data analysis - in this case
during the data preparation phase using spread sheets - can discredit
the work. Despite the increasing awareness, the adaption of
reproducibility or moving to other open scholarship/investigation
techniques is not yet achieved. This can be contrasted by emerging
practices like data journalism. One interesting example is buzzFeed's
article on the use of surveillance planes operated by law enforcement
and military for a variety of purposes. The authors were able to
demonstrate based on Open Data that these purposes range from tracking
drug trafficking to testing new surveillance technology
{[}\protect\hyperlink{ref-peteraldhous_2016}{7}{]}. The credibility of
the article comes from the fact that all artifacts are made available to
reproduce the findings or study the chosen approach and data
transformations made leading to the conclusions of the article.

Arguably there are limits to the access to underlying source data.
However, these limitations also offer an opportunity to address
reproducibility {[}\protect\hyperlink{ref-koelle_open_2017}{8}{]}. For
example, the data volume size might prohibit the further dissemination
of data, its collection is too resource intensive, or there exists
restrictive license agreements. Experience has shown that even in such
cases, a limited sample of the source data can be made available for
others to verify the initial data processing steps. For that purpose the
creation of an analytical data set as the output of the data collection,
gathering, and cleaning process stage can be understood as sufficient in
the sense of reproducible research. One key transparency aspect is the
documentation of the data preparation steps. The article Spies in the
Skies by Peter Aldhous and Charles Seife
{[}\protect\hyperlink{ref-peteraldhous_2016}{7}{]} is a good example for
a reproducible analysis of the US government's airborne surveillance
program. Though the license restrictions did not allow for the
publication of the underlying aircraft position data, the authors made
available the analytical data set and documented their analysis from the
data preparation through to the publication stage.

The reproducibility discussion focuses strongly on the recreation of
results by other researchers/analysts. A related problem is the
re-generating of already completed (and potentially published) results
with the - previously - available same input data. This is a fundamental
issue that can be summarised as ``can a future me re-do what I had done
in the past''? This problem space ranges across the complete data
analysis life-cyle, including archiving the orignially used input data
or the steps to generate the study data, documenting the data cleaning
and transformation steps including the parameters chosen, etc. By
definition such a capability is a precondition for reproduction or
replication by others. This implies that the analytical steps including
the steps to produce summaries or estimates, visualisations, and
associated conclusions are documented and archived. In order to enable
others to recreate the own findings, such steps and documentation would
need to be shareable. This consideration has a strong bearing on the
toolset chosen for this paper.

Some of the reasons why researchers/analysts do not publish data and
computational procedures that allow reproduction will be hard to change
{[}\protect\hyperlink{ref-daniel_nust_2018_2361304}{9}{]}. The
predominant technical reasons have been listed above. However, there are
also personal reasons such as the fear for floating ideas or loss of
competitive advantage. Other reasons deal with more technical problems,
e.g.~lack of capabilities, procedures.

Another incentive for the adoption of an open and reproducible research
practice can be seen in the wider recognition through increased
citations, media attention, discussion with the community and
collaborators, and ultimately funding opportunities. The recent years
have seen a strong use of research blogs and community platforms
supporting the publication of smaller steps or intermediate results. The
combination of web presentation and discussions offers a good basis for
an initial dialogue. Platforms like google scholar or researchgate.net
offer additional features for outreach and communication. This equally
applies for paper achives such as arxiv.org for pre-prints. Based on the
credibility crisis discussion a number of journals are now promoting -
or even requiring - the provision of analytical data in support of the
paper publication. There is also an increasing number of requirements
for national or European funded research with respect to data management
and Open Data access to the research.

Reproducibility is discussed in various domains. The following examples
further promote the use of open source software and/or data:

\begin{itemize}
\tightlist
\item
  archaeology: Marwick et al. (2017) focusses on computational
  reproducibility in archaeological research by providing basic
  principles. The case study of their implementation uses open source
  software (R/Rmarkdown/git approach)
  {[}\protect\hyperlink{ref-marwick_2017}{10}{]}. The paper advocates
  the publishing (analytical) data in support of the paper. This should
  not be an issue, even if there are constraints on the overall study
  data.
\item
  engineering / signal processing: Vandewalle (2009) discuss the
  motivation and approach to reproducibility in signal processing
  (i.e.~what, why, and how). Also here the case is made to open access
  to the code and data for validation purposes
  {[}\protect\hyperlink{ref-vandewalle_2009}{11}{]}.
\item
  operational ANS Performance monitoring: Spinielli et al. (2018)
  highlight the development and use of Open Flight Trajectories for
  reproducible ANS {[}\protect\hyperlink{ref-spinielli_2018}{5}{]}.
\end{itemize}

\begin{figure}[hbt]

{\centering \includegraphics[width=1\linewidth]{figures/reproducible-research-spectrum}

}

\caption{Reproducible research, Fig. 2 in
{[}\protect\hyperlink{ref-marwick_2017}{10}{]}.}\label{fig:reproducible-research}
\end{figure}

Across the literature, researchers have identified a ``reproducibility
spectrum'' ranging from the extreme, i.e.~publication only ( advertising
and marketing of the done work), to almost or even fully replicable
research. The work of Marwick et al. (2017)
{[}\protect\hyperlink{ref-marwick_2017}{10}{]} and Peng (2011)
{[}\protect\hyperlink{ref-Peng1226}{12}{]} uses the analogy of a
spectrum that can be mapped against the scoring of Vandewalle (2009)
describing the level of reproducibility on a 0 to 5 scale
{[}\protect\hyperlink{ref-marwick_2017}{10}{]}. In light of
{[}\protect\hyperlink{ref-leek_2015}{13}{]}, this paper conceptualises
reproducibility as the ``ability to recompute data analytic results
given an observed dataset and knowledge of the data analysis pipeline''
and accordingly ``replicability of a study is the chance that an
independent experiment targeting the same scientific question will
produce a consistent result''.

With Fig. \ref{fig:reproducible-research} and the discussion in this
section, this paper implements a fully reproducible workflow. This is
achieved by establishing this paper with open source software which
allows to combine text, code, and visualisations. The associated source
text can also easily be read with a simple text editor/viewer. From that
perspective, the text is fully inspectable. This includes the majority
of visualisations which are generated based on the data. Accordingly,
the parameters and settings chosen can be validated by an independent
researcher/analyst. Last but not least, the data for this paper is made
available in conjunction with this text as Open Data. This also includes
intermediate data artifacts which allows for the reproduction of the
complete work or portions of it. The conceptual approach and its
building blocks will be presented in more detail in the next section.

\section{Method / Approach}\label{method-approach}

\subsection{Conceptual Approach}\label{conceptual-approach}

\begin{figure}[hbt]

{\centering \includegraphics[width=1\linewidth]{figures/approach-stages}

}

\caption{Conceptual approach to reproducibility by PRU.}\label{fig:concept-approach}
\end{figure}

This paper builds on the conceptual approach presented in Fig.
\ref{fig:concept-approach} to implement replicable data analyses for
performance monitoring and associated research. On the data level this
approach abstracts three levels of data:

\begin{itemize}
\tightlist
\item
  source data
\item
  analytical data
\item
  results (and performance related data)
\end{itemize}

The source data for this paper comprises the PRU Performance Reference
Trajectory and operator reported data under the airport operator data
flow. This study also uses some aeronautical information for the use
case airports accessible through AIP or community maintained open data
archives (e.g.~http://openaip.net, http://ourairports.com). These data
flows are processed and stored in the EUROCONTROL data Warehouse. The
approach recognises that the full source data may or cannot be provided
for a variety of reasons. These include volume of the source data sets,
technical constraints allowing access to the source data storage, and
data sharing policies or use limitations. For the latter cases, PRU
promotes the sharing of a so-called representative sample. The idea is
that with the provision of the associated processing code, other
researchers/practitioners are able to reconstruct a part of the analytic
data - and subsequently part of the reported results - on the basis of
the provided source data set.

Analytical data refers to the data sets established based on the
aforementioned processing code. An essential defining characteristic of
analytic data is that it is the data basis on which the reported results
can be produced. The abstraction between source data and analytic data
allows for the sharing of such data as limitations are typically only
appliable to the source data. We follow the notion of
{[}\protect\hyperlink{ref-marwick_2017}{10}{]} by referring to the
provision of a representative sample of 8 days accompanying this paper
as source data.

As depicted in Fig. \ref{fig:reproducible-research} the provision of the
data, code, and text allows to move to the right part of the
reproducibility spectrum, away from marketing and final results to
credible, transparent, and challengeable analysis. Accordingly, the
results (or performance related data) level presents the final
abstraction.

\subsection{Toolset}\label{toolset}

This work builds on the RStudio tools for the R language
{[}\protect\hyperlink{ref-rcoreteam_2018}{14}{]} including Git (and the
web-based repository managers GitHub and GitLab) as underlying version
control system. The R language was originally developed within the
statistical community supporting the task of statistical reporting by
providing routines for the statistical computing and visualisation.
Being open source, the R community is actively engaging and sharing
related software packages to enhance the core functionality. Without
limiting the impact of other packages, the development of knitr
{[}\protect\hyperlink{ref-R-knitr}{15}{]} and RMarkdown
{[}\protect\hyperlink{ref-xie2018}{16}{]}, ggplot
{[}\protect\hyperlink{ref-wickham_2016}{17}{]} for visualisation, and
the so-called tidyverse packages and RStudio IDE
{[}\protect\hyperlink{ref-rstudioteam_2015}{18}{]} provide the basis for
an open source data analysis ecosystem.

A key feature for the implementation of the reproducibility paradigm is
the fact that RMarkdown supports the combination of text, coding, and
visualisations in a single document. The latter can be rendered into
various output formats, e.g.~classical paper report-styles (pdf, MS
Word), html web-pages, and slides (both classical Powerpoint style or
html-based presentations). The fact that outputs can be rendered to html
pages also supports the embedding of interactive graphics. This also
facilitates the communication and outreach amongst researchers/analysts
as standard web technology becomes a standard outlet format. It follows
that changes to the code (and associated visualisations) or text can be
readily updated by re-rendering the adapted report. This allows to
develop standard templates for recurring reports based on varying input
data, but it also serves the iterative development of larger reports
(and studies).

A minimal example of RMarkdown document, a plain-text file with the
conventional extension \texttt{.Rmd} is shown below:

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\OtherTok{---}
\FunctionTok{title:}\AttributeTok{ }\StringTok{"R Markdown example"}
\FunctionTok{author:}\AttributeTok{ }\StringTok{"John Doe"}
\FunctionTok{date:}\AttributeTok{ }\StringTok{"2019-02-17"}
\FunctionTok{output:}\AttributeTok{ html_document}
\OtherTok{---}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Some math }
\NormalTok{$\textbackslash{}frac\{1\}\{N\} \textbackslash{}sum_\{i=1\}^N (x_i -\textbackslash{}mu)^2$,}
\NormalTok{followed by numbered list:}

\NormalTok{1. }\FloatTok{*italics* and **bold** text}
\FloatTok{1. then bullets}
\FloatTok{    - simple, eh?}

\NormalTok{and a code chunk:}

\NormalTok{```\{r\}}
\NormalTok{library(ggplot2)}
\NormalTok{fit = lm(mpg ~ wt, data = mtcars)}
\NormalTok{b   = coef(fit)}
\NormalTok{ggplot(mtcars, aes(x = wt, y = mpg)) +}
\NormalTok{  geom_point() + }
\NormalTok{  geom_smooth(method = lm, se = TRUE)}
\NormalTok{```}

\NormalTok{The slope of the regression is }\BaseNTok{`r b[1]`}\NormalTok{.}
\NormalTok{```}
\end{Highlighting}
\end{Shaded}

\normalsize

\begin{figure}[htb]

{\centering \includegraphics[width=1\linewidth]{figures/rmd-example}

}

\caption{The output of a minimalist example of RMarkdown.}\label{fig:rmd-example}
\end{figure}

There are three components of an R Markdown document: the metadata, the
text and the code. The metadata, written between the pair of three
dashes \texttt{-\/-\/-}, helps in the production of output document
(title, type of document {[}web page, PDF, MS Word{]}, \ldots{}) The
text, which follows the metadata, is witten using Markdown, a simplified
formatting syntax. There are two types of (computer) code:

\begin{itemize}
\item
  A code chunk starts with three backticks and \texttt{r} indicates the
  language name,\footnote{R Markdown supports 50 other languages.} and
  ends with three backticks.
\item
  An inline R code expression starts with backtick r and ends with a
  backtick.
\end{itemize}

Code portions are executed when the document is rendered.

The result of rendering the R Markdown above is shown in Fig.
\ref{fig:rmd-example}.

A particular strong point is that the R/Rstudio ecosystem is open source
software. This includes also its extensibility with other packages.
There exists a wide variety of packages for data analytical tasks and
techniques. This mechanism can also be useful for the communication of
results. Researchers/analysts are able to provide their code and
supporting documentation as a package for other researchers to build on.
For example, throughout the last year, PRU has developed certain utility
functions for reoccurring tasks of the data analytical pipeline. It is
now planned to wrap these functions into a package saving other
researchers/analysts of recoding certain steps. For this paper, the
functions and data transformations are reported as part of the RMarkdown
and supporting R files. All these files are available via the project
webpage.

\subsection{Capabilities: Performance Reference
Trajectory}\label{capabilities-performance-reference-trajectory}

Throughout the recent year PRU developed the data analytical
capabilities to fuse correlated position reports and ADSB surveillance
data. This work has been reported in
{[}\protect\hyperlink{ref-koelle_open_2017}{8}{]} and
{[}\protect\hyperlink{ref-spinielli_2018}{5}{]}. The idea is to
establish a pan-European air situation picture openly available to
research and practitioners. Conceptually, the fused performance
reference trajectory will be a key milestone in combatting marketing
myths. For example, operational improvements can be directly assessed by
interested analysts with a view to validate the observable performance
improvements. This allows for the application of commonly accepted
methods and models to express performance benefits and - ultimately -
challenge related publications on deployments and achieved improvements.

End of 2018 the development project was closed and PRU is now
investigating options for deploying and maintaining the processing
modules on a cloud platform. The goal is to establish and operate the
underlying data processing capabilities throughout 2019 and launch the
regular provision of the performance reference trajectory as Open Data.
It is also planned to establish a network and complement the air
situation picture by deploying additional sensors or determining
non-ADSB derived positions as a community effort. As a mid-term goal
associated data bases and models for performance shall be established
and jointly maintained.

\section{Results}\label{results}

\subsection{Data Preparation}\label{data-preparation}

\begin{figure}[htb]

{\centering \includegraphics[width=1\linewidth]{figures/data-preparation-pipeline}

}

\caption{Data preparation pipeline.}\label{fig:data-preparation-pipeline}
\end{figure}

The raw data sources for the use cases of this paper have been processed
by a pipeline of R scripts. For example, the additional ASMA time
indicator and the anticipated further breakdown of the arrival phase can
be described by a set of tranistion points:

\begin{itemize}
\tightlist
\item
  segment I: entry into the ASMA area to entry into the
  holding/vectoring segment
\item
  segment II: holding/vectoring segment; for the use case reported
  holding stacks or point merge procedure
\item
  segment III: final approach alignment and approach
\end{itemize}

Figure \ref{fig:data-preparation-pipeline} shows how the transition
points (40 NM, entry/exit from holding/point-merge area, landing) for
the use case airports have been extracted. Due to the size of the daily
reference trajectory files, the pipeline operates on one date at the
time. Assuming the relevant RT flight and position files are downloaded
from GCP as described in
{[}\protect\hyperlink{ref-performancereviewunit_2019}{19}{]}, the steps
and the relevant scripts (available in the companion repository for this
paper) used are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  the script \texttt{R/extract-arrivals.R} extracts arrivals to the
  \texttt{APT} airport and generates:

  \begin{itemize}
  \tightlist
  \item
    a file with flight details
  \item
    a file with all the arrivals' positions
  \item
    a file with all arrivals' positions within 40 NM the airport's
    reference position (ARP)
  \end{itemize}
\item
  the script \texttt{R/extract-transition-points.R} extracts transition
  points in the terminal area and generates a file containing per each
  flight:

  \begin{itemize}
  \tightlist
  \item
    the first point within 40 NM, \texttt{P40}
  \item
    the first and last point within the holding/point-merge area,
    \texttt{PHOLD}
  \item
    an estimated landing point, \texttt{PLAND}
  \end{itemize}
\end{enumerate}

\begin{figure}[hbt]

{\centering \includegraphics[width=1\linewidth]{figures/eidw-arrival-chart-east}

}

\caption{Excerpt of Irish AIP showing point merge
arc layout East of Dublin Airport.}\label{fig:eidw-arrival-chart-east}
\end{figure}

This stage makes use of data extracted/extrapolated from the relevant
AIP, see for example Fig. \ref{fig:eidw-arrival-chart-east}, such as
ARP, runway name and threshold position/elevation, polygons for holding
stacks or point-merge areas. Specifically the polygons have been
heuristically defined and enlarged by 2NM in order to cater for the
spread of real flown trajectories, see the scripts
\texttt{R/egll-data.R} and \texttt{R/eidw-data.R}.

Each position mentioned above consist of at least

\begin{itemize}
\tightlist
\item
  flight ID
\item
  4D position (timestamp \(T\), longitude, latitude, altitude)
\item
  flown distance \(D\)
\item
  sequence number
\item
  distance from ARP
\end{itemize}

These allow to calculate for each flight

\begin{itemize}
\tightlist
\item
  the time spent before the holding/point-merge area,
  \(T_{PHOLD|first} - T_{P40}\)
\item
  the distance flown before the holding/point-merge area,
  \(D_{PHOLD | first} - D_{P40}\)
\item
  the time spent within the holding/point-merge area,
  \(T_{PHOLD | last} - T_{PHOLD|first}\)
\item
  the distance flown within the holding/point-merge area,
  \(D_{PHOLD | last} - D_{PHOLD | first}\)
\item
  the time spent out of the holding/point-merge area till landing,
  \(T_{PLAND} - T_{PHOLD | last}\)
\item
  the distance flown out of the holding/point-merge area till landing,
  \(D_{PLAND} - D_{PHOLD | last}\)
\end{itemize}

\subsection{Data Coverage}\label{data-coverage}

The previous sections described the data preparatory action. For this
paper, a reproducible data set for the analysis of arrival management
techniques at London Heathrow and Dublin airport have been extracted.
The data set comprises flights for the period of 1.-8. August 2017. It
can be inferred from the traffic mix at London Heathrow (major
international hub) that only a limited subset of flights is not equipped
with ADSB, while a slightly higher level of non-equipped flights is
expected at Dublin (higher share of intra-European flights).

\begin{figure}[hbt]

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/coverage-egll-1}

}

\caption{Number of arrivals in data sets, i.e.~1.-8. July
2017.}\label{fig:coverage-egll}
\end{figure}

Fig. \ref{fig:coverage-egll} shows the overlap of the data between the
arrivals extracted from the airport operator data flow (APDF) and the
reference trajectory (RTRJ) for London Heathrow (EGLL) and Dublin
(EIDW).

For the time horizon of this paper, the coverage of the reference
trajectory derived data is for Heathrow (EGLL) 100\% and for Dublin
97\%.

\subsection{\texorpdfstring{``replicate'' ASMA
(time-based)}{replicate ASMA (time-based)}}\label{replicate-asma-time-based}

The additional time in the arrival sequencing and metering area (ASMA)
is a globally recognised measure to assess flight efficiency during the
arrival phase. The ASMA metric calculates the additional time based on a
reference time which is determined for a family of flights sharing
similar characteristics. These are the entry sector into the arrival
airspace, weight turbulence category and engine type, and the actual
landing runway. For each flight \(i\) the additional ASMA time is

\[{add.~ASMA~time}_i = {actual~travel~time}_i - {reference~time}_i\]

The reference time \(i\) is determined for flights with similar arrival
characteristics (c.f.
\url{http://ansperformance.eu/references/definition/additional_asma_time.html}).
These include 1.) the arrival entry sector at 40NM from the airport, 2.)
the aircraft weigth turbulence category and engine type, and 3.) the
actual landing runway. The reference times are built for a longer time
horizon and extracted from the PRU data base as additional source data.

\begin{figure}[hbt]

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/ASMA-travel-egll-1}

}

\caption{Distribution of actual ASMA travel times for EGLL
(1.-8. July 2017).}\label{fig:ASMA-travel-egll}
\end{figure}

As an example, the distribution of the actual arrival transit times for
arrivals into Heathrow (EGLL) are shown in Fig.
\ref{fig:ASMA-travel-egll}. Heathrow operates holding stacks to
tactically optimise the arrival throughput. Dependent on the traffic
situation, aircraft may experience longer holding times (i.e.~long tails
of boxplots in Fig. \ref{fig:ASMA-travel-egll}).

Fig. \ref{fig:ASMA-compare} compares the additional ASMA time extracted
from the PRU performance dashboard, the trajectory based calculation
with the dashboard derived reference time, and a 20th percentile
approach purely based on the trajectory data.

\begin{figure}[hbt]

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/ASMA-compare-1}

}

\caption{Comparison of determined ASMA values, i.e.~1.-8. July
2017.}\label{fig:ASMA-compare}
\end{figure}

Fig. \ref{fig:ASMA-compare} shows the variation of the timebased
additional ASMA times determined for our sample. APDFNM refers to the
variant used for the EUROCONTROL Performance Review System. The
reference time is build on the basis of a one-year sammple. Applying
this reference time to the trajectory based travel times (i.e.~TRJ-REF),
there is a strong increase of the additional ASMA time for Heathrow
(EGLL), while in the case of Dublin (EIDW) reasonable decrease takes
place. Calculating the additional ASMA time exclusively with the
trajectory sample and applying a 20th percentile approach (TRJ-20PCT),
shows a significant drop in EGLL and a moderate increase at Dublin. The
high variation for Heathrow shows that the additional ASMA time is
strongly dependent on the time horizon. The pure trajectory bases
measure - by definition - includes the effect of the operational concept
(stack holdings) and distorts the measure. In the case of Dublin it can
be argued that the TRJ-20PCT measure reflects the time based ASMA
measure suffiently.

\subsection{Distance-based ASMA}\label{distance-based-asma}

conceptually, measuring flight inefficiency in terms of flight time
accounts for the airspace user perspective. The previous section has
also shown the high dependency of the time based measure on the sample
size or reference time attribution. Flight time is linked to engine time
and accordingly, assuming maneuvering for operational separation and
sequencing reasons, fuel burn. From a service provision perspective,
efficiency relates to the capability to provide the least distance flown
within the ASMA area. In that respect, the data analytical capability to
determine an associated additional distance flown metric addresses an
open gap.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/ASMA-time-dist-1}

}

\caption{Relationship travel time and distance flown,
i.e.~1.-8. July 2017.}\label{fig:ASMA-time-dist}
\end{figure}

Fig. \ref{fig:ASMA-time-dist} shows the relationship between the actual
distance flown and the actual time flown. The variation of travel times
is influenced by varying airspeeds or weather phenomena (e.g.~wind).
This paper applies a simple 15th percentile approach to the distance
flown per arrival sector, aircraft type/engine, and landing runway.

\begin{table}[t]

\caption{\label{tab:tbl-asma-dist}Average additional distance flown (15th pct)}
\centering
\begin{tabular}{l|r|r|r}
\hline
AIRPORT & TOT\_ADD\_DIST & N\_ARRS & AVG\_ADD\_DIST\\
\hline
EGLL & 117037.79 & 5249 & 22.30\\
\hline
EIDW & 33079.22 & 2634 & 12.56\\
\hline
\end{tabular}
\end{table}

Tab. \ref{tab:tbl-asma-dist} lists the determined average additional
distance at the study airports.

\subsection{Arrival Management
Practices}\label{arrival-management-practices}

For an initial drill down, this paper uses arrival practices at London
Heathrow and Dublin. The approach to managing the arrival flow differs
significantly between both airports. Conceptually, London Heathrow
employs the ``pressure on the runway'' concept with a focus on achieving
and balancing a continuously high level of runway throughput
utilisation. This is achieved by controlling the demand through 4
arrival stacks that allow for the tactical sequencing on the last
portion of the approach. Dublin is one of the airports that pioneered
the implementation of ``point merge''. Point merge aims at replacing
circular holdings by longitudal holdings to establish the sequence based
on predefined legs.

From a performance perspective it is therefore interesting to identify
and measure the impact of the operational concept on the overall flight
efficiency during the arrival phase of flight.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/egll-trajectory-example-1}

}

\caption{EGLL - long trajectory flown on 1. July
2017.}\label{fig:egll-trajectory-example}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/eidw-trajectory-example-1}

}

\caption{EIDW - long trajectory flown on 1. July
2017.}\label{fig:eidw-trajectory-example}
\end{figure}

Fig. \ref{fig:egll-trajectory-example} and Fig.
\ref{fig:eidw-trajectory-example} show examples for long distances flown
on 1. July 2017. The trajectory data allows for the identification of
the holdings and/or sequencing area. This allows to separate the effect
of the operational concepts.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/distance-flown-segment-1}

}

\caption{Actual distances flown for each segment of
the arrival phase (i.e.~EGLL, 1.-8. July 2017).}\label{fig:distance-flown-segment}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/distance-flown-segment-eidw-1}

}

\caption{Actual distances flown for each
segment of the arrival phase (i.e.~EIDW, 1.-8. July 2017).}\label{fig:distance-flown-segment-eidw}
\end{figure}

Fig. \ref{fig:distance-flown-segment} depicts the flown distances per
segment of the arrival phase. In the case of London Heathrow the
classical stack based holding can be seen in the significant variations
of the actual distance flown per arrival direction. Only limited
sequencing takes place during the first segment (arrival within 40 NM
and the stacks). Fig. \ref{fig:distance-flown-segment-eidw} shows
different footprint. It follows from the distribution of the initial
approach segment that this portion is already used for sequencing and
synchronisation by Dublin approach. The middle segment shows a wider
variation and less uniform distribution than the approach at London
Heathrow.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/eidw-saturation-example-1}

}

\caption{EIDW - arrival throughput saturation
(1.-8. July 2017).}\label{fig:eidw-saturation-example}
\end{figure}

Fig. \ref{fig:eidw-saturation-example} depicts the variation of the
arrival segment throughput at Dublin (EIDW). The median for every set of
flights in the ASMA area versus flights landing is highlighted. The
behavour is a classical saturation curve, i.e.~up to an arrival bunching
of 10-11 flights per 20 min (30-33 flights per hour), the point merge
procedure seems to be able to support the tactical absorption and
sequencing. It is noteworthy that the saturation curve falls for higher
arrival throughputs. As no mearurements for 12 through 14 landings had
been made in the sample, this behaviour could be caused by the chosen
time horizon.

\section{Conclusions}\label{conclusions}

The work reported establishes a fully reproducible paper for ANS
performance in Europe. It is the one of the first attempts to address
the reproducibility crisis within ANS performance by providing the
artifacts that allow an independent researcher or professional to
reproduce the results and verify upstream processing steps.

This paper follows up on previous work
{[}\protect\hyperlink{ref-spinielli_2018}{5}{]},
{[}\protect\hyperlink{ref-koelle_open_2017}{8}{]},
{[}\protect\hyperlink{ref-spinielli_2017}{20}{]} and forms part of the
roll-out of the performance reference trajectory by the PRU. As such the
paper demonstrates the feasibility to the approach, the data analytical
capabilities and preprocessing, and the reporting via the R/RStudio
ecosystem. This represents an essential milestone moving towards the
envisaged target environment. However, further work is required to
establish the reproducible environment for ANS Performance.

To overcome these limitations, PRU will be looking into deploying the
trajectory processing on a cloud-based platform including the regular
provision of the identified analytical data levels.

The move to open data for ANS Performance is a \emph{giant leap} in
terms of transparency. This will foster a new level of cooperation and
discussion amongst researchers and professionals interested in this
topic. It removes the burden to acquire and preprocess large-scale
surveillance data and associated aircraft data or aeronautical
information. PRU envisages a strong collaboration through a network of
research and practitioners. The departure point for this development was
to address the lack and hurdle of verifying results. A further advantage
is that novel applications, methods and algorithms (e.g.~machine
learning), can be applied to a harmonised dataset. It is hoped that
development efforts can accelerate while offering a platform for
improvements to existing processing methods and algorithms.

The major application of the paper results will inform the performance
reference trajectory roll-out project. This goes in hand with the
iterative move of the operational performance measures to trajectory
derived data. This can take on the form of quality assurance measures
(e.g.~verifying the order of magnitude, complementing missing reported
data), but also making the reference trajectory the primary means of
data collection. The latter will support the further development of the
operational performance measures responding to the ICAO GANP.

This paper documents a key milestone in the project. As such the future
work will address:

\begin{itemize}
\tightlist
\item
  roll-out and establishment of the performance reference trajectory:
  This includes the organisational and technical work to deploy the
  processing modules and arrange for a regular processing of the source
  data and its provision as Open Data.
\item
  establishing the current performance reporting based on the
  trajectory: Since more than 2 years PRU addresses the need for Open
  Data by providing monthly updates for the key performance indicators
  used under the EUROCONTROL Performance Review System and the SES
  Performance Scheme. Underlying analytic data is made available via
  \url{http://ansperformance.eu}. As part of the roll-out, it is
  foreseen to move associated indicators to a trajectory derived or at
  least verified process and expand on the analytical data level.
\item
  establishment of a network: The underlying idea is that various levels
  of future work can be orchestrated in close collaboration with
  interested researchers and professionals. This can span work
  addressing the improvement of the actual data processing (e.g.~less
  resource intensive computing) to the further development and
  validation of novel techniques for measuring operational performance
  (e.g.~application of machine learning, time series analysis).
\end{itemize}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\vspace{6pt} \setlength{\parskip}{0pt} \footnotesize
\setlength{\parindent}{0cm}

\hypertarget{refs}{}
\hypertarget{ref-kolata_2011}{}
{[}1{]} G. Kolata, ``How a New Hope in Cancer Fell Apart,'' \emph{The
New York Times: Health}, 07-Jul-2011.

\hypertarget{ref-ritchie_2012}{}
{[}2{]} S. J. Ritchie, R. Wiseman, and C. C. French, ``Failing the
future: Three unsuccessful attempts to replicate Bem's `Retroactive
Facilitation of Recall'Effect,'' \emph{PloS one}, vol. 7, no. 3, p.
e33423, 2012.

\hypertarget{ref-fraser_etal_2018}{}
{[}3{]} H. Fraser, T. Parker, S. Nakagawa, A. Barnett, and F. Fidler,
``Questionable research practices in ecology and evolution,'' \emph{PloS
one}, vol. 13, no. 7, p. e0200303, 2018.

\hypertarget{ref-fidler_wilcox_2018}{}
{[}4{]} F. Fidler and J. Wilcox, ``Reproducibility of Scientific
Results,'' in \emph{The Stanford Encyclopedia of Philosophy}, Winter
2018., E. N. Zalta, Ed. Metaphysics Research Lab, Stanford University,
2018.

\hypertarget{ref-spinielli_2018}{}
{[}5{]} E. Spinielli, R. Koelle, K. Barker, and N. Korbey, ``Open Flight
Trajectories for Reproducible ANS Performance Review,'' presented at the
SESAR Innovation Days 2018, 2018, p. 8.

\hypertarget{ref-gandrud_2015}{}
{[}6{]} C. Gandrud, \emph{Reproducible Research with R and R Studio,
Second Edition (Chapman \& Hall/CRC The R Series)}. Routledge, 2015.

\hypertarget{ref-peteraldhous_2016}{}
{[}7{]} C. S. Peter Aldhous, ``Spies in the Skies,'' \emph{See Maps
Showing Where FBI Planes Are Watching From Above}, 06-Apr-2016.
{[}Online{]}. Available:
\url{https://www.buzzfeed.com/peteraldhous/spies-in-the-skies}.

\hypertarget{ref-koelle_open_2017}{}
{[}8{]} R. Koelle, ``Open Source Software and Crowd Sourced Data for
Operational Performance Analysis,'' presented at the USA/Europe ATM R\&D
Seminar, 2017.

\hypertarget{ref-daniel_nust_2018_2361304}{}
{[}9{]} D. Nüst, M. Schutzeichel, and M. Konkol, ``Opening Reproducible
Research: A research project website and blog,'' Nov. 2018.

\hypertarget{ref-marwick_2017}{}
{[}10{]} B. Marwick, ``Open Science in Archaeology,'' \emph{Open Science
Framework}, Jan. 2017.

\hypertarget{ref-vandewalle_2009}{}
{[}11{]} P. Vandewalle, J. Kovacevic, and M. Vetterli, ``Reproducible
research in signal processing,'' \emph{IEEE Signal Processing Magazine},
vol. 26, no. 3, pp. 37--47, May 2009.

\hypertarget{ref-Peng1226}{}
{[}12{]} R. D. Peng, ``Reproducible Research in Computational Science,''
\emph{Science}, vol. 334, no. 6060, pp. 1226--1227, 2011.

\hypertarget{ref-leek_2015}{}
{[}13{]} J. T. Leek and R. D. Peng, ``Opinion: Reproducible research can
still be wrong: Adopting a prevention approach,'' \emph{Proc Natl Acad
Sci USA}, vol. 112, no. 6, p. 1645, Feb. 2015.

\hypertarget{ref-rcoreteam_2018}{}
{[}14{]} R Core Team, \emph{R: A Language and Environment for
Statistical Computing}. Vienna, Austria: R Foundation for Statistical
Computing, 2018.

\hypertarget{ref-R-knitr}{}
{[}15{]} Y. Xie, \emph{Knitr: A General-Purpose Package for Dynamic
Report Generation in R}. 2018.

\hypertarget{ref-xie2018}{}
{[}16{]} Y. Xie, J. Allaire, and G. Grolemund, \emph{R Markdown: The
Definitive Guide}. Boca Raton, Florida: Chapman and Hall/CRC, 2018.

\hypertarget{ref-wickham_2016}{}
{[}17{]} H. Wickham, \emph{Ggplot2: Elegant Graphics for Data Analysis}.
Springer-Verlag New York, 2016.

\hypertarget{ref-rstudioteam_2015}{}
{[}18{]} RStudio Team, \emph{RStudio: Integrated Development Environment
for R}. Boston, MA: RStudio, Inc., 2015.

\hypertarget{ref-performancereviewunit_2019}{}
{[}19{]} Performance Review Unit, ``Reference trajectories for
Performance Review of European ANS,'' 2019. {[}Online{]}. Available:
\url{https://doi.org/10.5281/zenodo.2566844}.

\hypertarget{ref-spinielli_2017}{}
{[}20{]} E. Spinielli, R. Koelle, M. Zanin, and S. Belkoura, ``Initial
Implementation of Reference Trajectories for Performance Review,''
\emph{SESAR Innovation Days}, no. 7, p. 8, 2017.

\setlength{\parskip}{6pt}

\vspace*{1\baselineskip}
\hypertarget{biographies}{%
\section*{Authors' Biographies}\label{biographies}}
\addcontentsline{toc}{section}{Authors' Biographies}

\vspace*{-3\baselineskip}

\begin{IEEEbiographynophoto}{\normalsize Enrico Spinielli}\normalsize
is the Data Manager of the Performance Review Unit at EUROCONTROL, Brussels.
He holds a Master Degree in Electronic Engineering from the University of Pisa, Italy.
His intellectual interests include data analytics, data visualization and
map projections.
\end{IEEEbiographynophoto}

\vspace*{-2.5\baselineskip}

\begin{IEEEbiographynophoto}{\normalsize Rainer Koelle}\normalsize
heads the Safety and Quality of Service section
of the Performance Review Unit at EUROCONTROL, Brussels.
He holds a PhD from the University of Lancaster, 2013, in Communication Systems.
His research interests include data analytics for operational performance analysis.
\end{IEEEbiographynophoto}

\vfill

\end{document}