### 2019-02-18 analysis/data.Rmd
fixed few typos and markdown

---
title: "Data"
output:
  workflowr::wflow_html:
    toc: false
---

# Data Files

We have two directories for data

1. `data-raw`: it contains the raw source data. For this project these are the Reference Trajectory
   files as from Google Cloud Platform. See https://github.com/euctrl-pru/reftrj .
   Due to their big size these are not duplicated/version controlled on this repo.
1. `data`: here are all the files necessary for the analysis as performed in the paper.

## From `data-raw` to `data`

The raw data filenames are as follows:

* flight information, i.e. `cpr_fr24_flights_2017-08-01.csv.bz2`
* positions data, i.e. `mas_05_cpr_fr24_synth_positions_2017-08-01.csv.bz2`

PRU's [Open Reference Trajectory dataset](https://github.com/euctrl-pru/reftrj)
makes available the files needed for the analysis in this project;
you can retrieve them either by saving the relevant linked
flight and position files from your browser or by using Google `gsutil` as follows:

```{bash ,eval=FALSE}
# enable relevant Python virtual environment (Python 2) if you have one, and you should.
$ pyenv activate anaconda2-4.2.0
$ cd data-raw
$ gsutil cp \
    gs://pru-trajectories/sources/merged/overnight_cpr_fr24/raw_cpr_fr24_positions_2017-08-08.csv.bz2 .
```

As explained in the paper, you can use `R/extract-arrivals.R` to produce CSV files
containing only the flights and positions for your airport of initerest.
You should edit the relevant configuration values at the beginning of the script before launching it.
You will get:

* Flights details for arrivals at `<airport>` on `<date>`
  are in a CSV file `<airport>_flt_rt_<date>.csv`.
  For example `egll_flt_rt_2017-08-01.csv` is for arrivals to
  London Heathrow Airport (ICAO's id EGLL) on Aug, 1 2017.
  Key data items:
    * `flight_id`     unique flight identification
    * `callsign`      callsign flight identification of flight
    * `aircraft_reg`  aircraft registration
    * `aircraft_type` aircraft type, ICAO coding (e.g. A388, B772)
    * `adep` / `ades`   aerodrome of departure / destination

* Position reports for arrival at `<airport>` on `<date>`
  are in a CSV file `<airport>_pos_rt_<date>.csv`.
  For example `eidw_pos_rt_2017-08-01.csv` is for arrivals to
  Dublin Airport (ICAO's id EIDW) on Aug, 1 2017.

The focus of this project is for the terminal area, so we have further filtered the data
and kept the last 40 NM from the airport reference position (ARP).

The position reports within 40 nautical miles for arrival at
`<airport>` on `<date>`  are in a CSV file 
`<airport>_pos_rt_40NM_<date>.csv`.
These are a subset from `<airport>_pos_rt_<date>.csv`.
For example `eidw_pos_rt_40NM_2017-08-02.csv` is for arrivals'
positions within 40 nautical miles to Dublin Airport
(ICAO's id EIDW) on Aug, 2 2017.
Key data items:

  * `flight_id`     unique flight identification
  * `timestamp`     position report time
  * `latitude`      latitude coordinate (decimal notation)
  * `longitude`     longitude coordinate (decimal notation)
  * `on_ground`     on-ground / aloft flag
  * `distance`      flown distance
  * `distance_arp`  distance position to aerodrome reference point

Further we have extracted the transition points from the holding/point merge areas
using the script `R/extract-transition-points.R` which produces
the files `<airport>_transition_points_<date>.csv`.

## Version controlled data

We version control only the following data sets in `data`

* `<airport>_flt_rt_<date>.csv`
* `<airport>_pos_rt_40NM_<date>.csv`
* `<airport>_transition_points_<date>.csv`

No files are version controlled in `data-raw`.