### 2019-09-26 analysis/setup.Rmd
Improved installation instructions for rticles

---
title: "Setup"
output:
  workflowr::wflow_html:
    toc: false
---

This RMarkdown describes the setup of the R/RStudio environment needed to reproduce the paper.

## Reproduce the paper

The analysis and production of this paper make use of a number of R packages, which need to be
installed and functioning:

* `tinytex`
  ```{r, eval=FALSE}
  devtools::install_github('yihui/tinytex')
  tinytex::install_tinytex()
  ```

* `rticles`: from [`rstudio`](https://github.com/rstudio/rticles) if
  [PR #247](https://github.com/rstudio/rticles/pull/247) has been merged
  ```{r, eval=FALSE}
  install.packages("rticles")
  ```
  otherwise directly from [PR #247](https://github.com/rstudio/rticles/pull/247)
  ```{r, eval=FALSE}
  remotes::install_github("rstudio/rticles#247")
  ```

* `tidyverse` and `sf`
  ```{r, eval=FALSE}
  install.packages(c("tidyverse", "sf"))
  ```

* install IEEE Transaction LaTeX style:
    ```{r, eval=FALSE}
    tinytex::tlmgr_install("IEEEtran")
    ```

* in order to overcome the missing small caps fonts (this needs to be done once):
  1. uncomment `\usepackage{times}` in `header.tex`
  2. knit `paper.Rmd`
  3. comment again `\usepackage{times}` in `header.tex`

You should now be in the position to knit `paper.Rmd` and obtain `paper.pdf`.

## Issues

  * References on second and subsequent lines not correctly indented.