### 2019-09-26 analysis/website.Rmd
Added info on paper and website production

---
title: "Website"
output:
  workflowr::wflow_html:
    toc: false
---

## Build the companion website

The companion website to this project is generated by the `workflowr` package as follows:

```{r website-build,eval=FALSE}
wflow_build()
```

This will 
1. generate the relevant HTML pages in `docs` out of the R markdown files in the `analysis`
   directory.
1. preview the website in the Viewer panel in RStudio

The `docs` directory is automatically served by GitHub, so the website will be browsable on
the Internet at https://euctrl-pru.github.io/reproducible-ans-performance-paper/.

## Publish the website

Publishing and versioning the project website us accomplished via `wflow_publish()`.
This will build and version control the HTML pages.